### 2018-05-04 docs/lc-emulsion.tex

\subsubsection{Control of the isotropic/nematic transition}
The function $\gamma(\phi)$ allows control of the liquid crystal
ordering as the composition changes. An appropriate choice of function
for the case of $\phi^\star = 1$ is
\begin{equation}
\gamma(\phi)=\gamma_0+\delta\left(1+\phi\right)
\end{equation}
so that $\gamma(\phi=-1)=\gamma_0$, and $\gamma(\phi=1)=\gamma_0+2\delta$.
We should choose $\gamma_0$ and $\delta$ such that $\gamma(\phi=-1)$ 
is smaller than 2.7 (so that for $\phi=-1$ the fluid is isotropic)
and $\gamma(\phi=+1)$ is larger than 2.7 (so that when $\phi=1$ the
liquid crystal is in the ordered phase, nematic or cholesteric etc.).
Experience suggests that for example 
$\gamma_0=2.5$ and $\delta=0.25$ is a suitable choice.

@@

In the
case that $\gamma(\phi) \sim \delta\phi$ we generate terms
\begin{equation}
\mu_{\rm c} + \mu_{\rm lc} = A\phi + B\phi^3 -\kappa \partial_{\gamma}\partial_{\gamma}\phi
- \textstyle{\frac{1}{6}} A_0 \delta Q_{\alpha\beta}^2
- \textstyle{\frac{1}{3}} A_0 \delta
                          Q_{\alpha\beta} Q_{\beta\pi} Q_{\pi\alpha}
+ \textstyle{\frac{1}{4}} A_0 \delta (Q_{\alpha\beta}^2)^2
\end{equation}
from $f_{\rm c}(\phi)$ and $f_{\rm lc}(Q_{\alpha\beta})$.
The interfacial or anchoring contribution to the chemical potential is
\begin{equation} 
\mu_{\rm anchoring}=-W \left(
 \partial_{\alpha}\phi\partial_{\beta}Q_{\alpha\beta}
+\partial_{\beta}\phi\partial_{\alpha}Q_{\alpha\beta}\right)
- 2W Q_{\alpha\beta} \partial_\alpha\partial_\beta\phi.
\end{equation}

@@

It is convenient \cite{sulaiman2006} to rewrite a divergence of the
stress  $-\partial_\beta P_{\alpha\beta}$ with the stress defined in
Equation~\ref{eq-lc-emulsion-pab}. This force can be expanded in full as

\begin{equation}
\partial_{\alpha}\left(\phi\mu - F\right)
+\partial_{\beta} \left(\frac{\delta F}{\delta \left(\partial_{\beta}\phi\right)}\right)\partial_{\alpha}\phi
+\frac{\delta F}{\delta \left(\partial_{\beta}\phi\right)}\partial_{\beta}\left(\partial_{\alpha}\phi\right)
+\partial_{\beta}\left(\frac{\delta F}{\delta Q_{\pi\nu,\beta}}\right)
 \partial_{\alpha} Q_{\pi\nu}
+\frac{\delta F}{\delta Q_{\pi\nu,\beta}} \partial_{\beta}\left(\partial_{\alpha} Q_{\pi\nu}\right).
\nonumber
\end{equation}
With some manipulation and using $$\frac{\partial F}{\partial (\partial_\alpha\partial_\beta\phi)} = \frac{\partial F}{\partial (\partial_\alpha Q_{\mu\nu,\beta})}=0$$
this can be recast as 
\begin{eqnarray}
-\partial_\beta P_{\alpha\beta} &=& -\left( \frac{\partial F}{\partial Q_{\pi\nu}} 
       - \partial_{\beta}\frac{\partial F}{\partial Q_{\pi\nu,\beta}}
 \right) \partial_{\alpha}Q_{\pi\nu}
- \left(\frac{\partial F}{\partial\phi} 
- \partial_{\beta} \frac{\partial F}{\partial \left(\partial_{\beta}\phi\right)}\right)\partial_{\alpha}\phi + \partial_{\alpha}\left(\mu\phi\right)\nonumber\\
&=& -\left(\frac{\delta F}{\delta Q_{\pi\nu}}\right) \partial_{\alpha}Q_{\pi\nu} - \left(\frac{\delta F}{\delta \phi}\right) \partial_{\alpha} \phi + \partial_{\alpha}\left(\mu\phi\right)\nonumber\\ 
&=& H_{\pi\nu}\; \partial_{\alpha} Q_{\pi\nu} + \phi \;\partial_{\alpha}\mu,
\label{eq-lc-emulsion-hgradq}
\end{eqnarray}
which is a local force on the fluid.
The total force on the fluid locally is the divergence of the stress,
where the total stress is $\Pi_{\alpha\beta}$
given by equation~(\ref{eq-lc-emulsion-stress}).
It is sometimes useful to compute the various contributions to the
force in different ways. The contribution from the symmetric and
antisymmetric stress $\sigma_{\alpha\beta}$ and  $\tau_{\alpha\beta}$
can be calculated numerically as a divergence. The contribution from
$P_{\alpha\beta}$ can be computed numerically as a divergence, or via the
analytical form of the divergence Equation~\ref{eq-lc-emulsion-hgradq}.
The  last option is
at the expense of global conservation of momentum.

>>>>

\subsubsection{Control of the isotropic/nematic transition}
The function $\gamma(\phi)$ allows control of the liquid crystal
ordering as the composition changes. An appropriate choice of function
for the case of $\phi^\star = 1$ is
\begin{equation}
\gamma(\phi)=\gamma_0+\Delta\left(1+\phi\right)
\end{equation}
so that $\gamma(\phi=-1)=\gamma_0$, and $\gamma(\phi=1)=\gamma_0+2\Delta$.
We should choose $\gamma_0$ and $\Delta$ such that $\gamma(\phi=-1)$ 
is smaller than 2.7 (so that for $\phi=-1$ the fluid is isotropic)
and $\gamma(\phi=+1)$ is larger than 2.7 (so that when $\phi=1$ the
liquid crystal is in the ordered phase, nematic or cholesteric etc.).
Experience suggests that for example 
$\gamma_0=2.5$ and $\Delta=0.25$ is a suitable choice.

@@

In the
case that $\gamma(\phi) \sim \Delta\cdot\phi$ we generate terms
\begin{equation}
\mu_{\rm c} + \mu_{\rm lc} = A\phi + B\phi^3 -\kappa \partial_{\gamma}\partial_{\gamma}\phi
- \textstyle{\frac{1}{6}} A_0 \Delta Q_{\alpha\beta}^2
- \textstyle{\frac{1}{3}} A_0 \Delta
                          Q_{\alpha\beta} Q_{\beta\pi} Q_{\pi\alpha}
+ \textstyle{\frac{1}{4}} A_0 \Delta (Q_{\alpha\beta}^2)^2
\end{equation}
from $f_{\rm c}(\phi)$ and $f_{\rm lc}(Q_{\alpha\beta})$.
The interfacial or anchoring contribution to the chemical potential is
\begin{equation} 
\mu_{\rm anchoring}=-W \left(
 \partial_{\alpha}\phi\partial_{\beta}Q_{\alpha\beta}
+\partial_{\beta}\phi\partial_{\alpha}Q_{\alpha\beta}\right)
- 2W Q_{\alpha\beta} \partial_\alpha\partial_\beta\phi.
\end{equation}

@@

It is convenient \cite{sulaiman2006} to rewrite a divergence of the
stress  $-\partial_\beta P_{\alpha\beta}$ with the stress defined in
Equation~\ref{eq-lc-emulsion-pab}. This force can be expanded in full as
\begin{equation}
\partial_{\alpha}\left(\phi\mu - F\right)
+\partial_{\beta} \left(\frac{\delta F}{\delta \left(\partial_{\beta}\phi\right)}\right)\partial_{\alpha}\phi
+\frac{\delta F}{\delta \left(\partial_{\beta}\phi\right)}\partial_{\beta}\left(\partial_{\alpha}\phi\right)
+\partial_{\beta}\left(\frac{\delta F}{\delta Q_{\pi\nu,\beta}}\right)
 \partial_{\alpha} Q_{\pi\nu}
+\frac{\delta F}{\delta Q_{\pi\nu,\beta}} \partial_{\beta}\left(\partial_{\alpha} Q_{\pi\nu}\right).
\label{eq-lc-emulsion-force}
\end{equation}
Applying the chain rule on the first bracket in Equation \ref{eq-lc-emulsion-force} leads to
\begin{equation}
\partial_{\alpha}(\phi\mu) 
- \frac{\partial F}{\partial \phi} \partial_\alpha \phi
- \frac{\partial F}{\partial(\partial_\beta \phi)} \partial_\alpha (\partial_\beta\phi)
- \frac{\partial F}{\partial Q_{\pi\nu}} \partial_\alpha Q_{\pi\nu}  
- \frac{\partial F}{\partial Q_{\pi\nu,\beta}} \partial_\alpha Q_{\pi\nu,\beta}\nonumber.
\end{equation}
The expansion of the functional derivatives in Equation \ref{eq-lc-emulsion-force} gives
\begin{eqnarray}
&&\partial_{\beta} \left(\frac{\partial F}{\partial (\partial_{\beta}\phi)} - \partial_\gamma \frac{\partial F}{\partial (\partial_\gamma\partial_\beta\phi)}\right) \partial_{\alpha}\phi
+\left(\frac{\partial F}{\partial (\partial_{\beta}\phi)} - \partial_\gamma \frac{\partial F}{\partial (\partial_\gamma\partial_\beta\phi)}\right) \partial_\beta\left(\partial_{\alpha}\phi\right)\nonumber\\
&&\partial_{\beta} \left(\frac{\partial F}{\partial Q_{\pi\nu,\beta}} - \partial_\gamma \frac{\partial F}{\partial (\partial_\gamma Q_{\pi\nu,\beta)}}\right) \partial_{\alpha}Q_{\pi\nu}
+ \left(\frac{\partial F}{\partial Q_{\pi\nu,\beta}} - \partial_\gamma \frac{\partial F}{\partial (\partial_\gamma Q_{\pi\nu,\beta)}}\right) \partial_\beta Q_{\pi\nu,\alpha}\nonumber.
\end{eqnarray}
Using the symmetry of partial derivatives and $$\frac{\partial F}{\partial (\partial_\gamma\partial_\beta\phi)} = \frac{\partial F}{\partial (\partial_\gamma Q_{\mu\nu,\beta})}=0$$
this can be recast with some manipulation as 
\begin{eqnarray}
-\partial_\beta P_{\alpha\beta} &=& -\left( \frac{\partial F}{\partial Q_{\pi\nu}} 
       - \partial_{\beta}\frac{\partial F}{\partial Q_{\pi\nu,\beta}}
 \right) \partial_{\alpha}Q_{\pi\nu}
- \left(\frac{\partial F}{\partial\phi} 
- \partial_{\beta} \frac{\partial F}{\partial \left(\partial_{\beta}\phi\right)}\right)\partial_{\alpha}\phi + \partial_{\alpha}\left(\phi\mu\right)\nonumber\\
&=& -\left(\frac{\delta F}{\delta Q_{\pi\nu}}\right) \partial_{\alpha}Q_{\pi\nu} - \left(\frac{\delta F}{\delta \phi}\right) \partial_{\alpha} \phi + \partial_{\alpha}\left(\phi\mu\right)\nonumber\\ 
&=& H_{\pi\nu}\; \partial_{\alpha} Q_{\pi\nu} + \phi \;\partial_{\alpha}\mu,
\label{eq-lc-emulsion-hgradq}
\end{eqnarray}
which is a local force on the fluid.
The total force on the fluid locally is the divergence of the stress,
where the total stress is $\Pi_{\alpha\beta}$
given by equation~(\ref{eq-lc-emulsion-stress}).
It is sometimes useful to compute the various contributions to the
force in different ways. The contribution from the symmetric and
antisymmetric stress $\sigma_{\alpha\beta}$ and  $\tau_{\alpha\beta}$
can be calculated numerically as a divergence. The contribution from
$P_{\alpha\beta}$ can be computed numerically as a divergence, or via the
analytical form of the divergence Equation~\ref{eq-lc-emulsion-hgradq}.
The  last option is
at the expense of global conservation of momentum.

### 2018-05-04 docs/references.tex
Trunk add LC emulsion document and relevant changes
git-svn-id: https://ccpforge.cse.rl.ac.uk/svn/ludwig/trunk@3079 40876441-3496-0410-a613-a2fca35b2ae2

\bibitem{swift1996}
M.R. Swift, E. Orlandini, W.R. Osborn, and J.M. Yeomans,
Lattice Boltzmann simulations of liquid-gas and binary fluid systems,
\textit{Phys. Rev. E} \textbf{54} 5041--5052 (1996).

@@

@@

\bibitem{Lyklema} J. Lyklema {\em Fundamentals of Interface and Colloid Science} Academic Press     (1995).
\bibitem{Mafe} S. Maf\'e, J.A. Manzanares, J. Pellicer, {\textit J. Electroanal. Chem.} {\textbf 241}, 5    7-77 (1988).
\bibitem{Capuani} F. Capuani, I. Pagonabarraga, D. Frenkel, {\textit J. Chem. Phys.} {\textbf 121}, 973-    986 (2004).
\bibitem{Rotenberg} B. Rotenberg, I. Pagonabarraga, D. Frenkel, {\textit Farad. Discuss.} {\textbf 144},     223-243 (2010).
\bibitem{Landau-ED} L.D. Landau, E.M. Lifshitz, {\textit Electrodynamics of Continuous Media}, \S 15    , 2nd ed., Pergamon Press, Oxford, UK (1984).
\bibitem{Melcher} J.R. Melcher, {\textit Continuum Electromechanics}, \S 3.10, MIT Press, Cambridge,     MA, USA (1981).\\
downloadable from:\\
\url{http://ocw.mit.edu/ans7870/resources/melcher/resized/cem_811.pdf}
\bibitem{Landau-EL} L.D. Landau, E.M. Lifshitz, {\textit Theory of Elasticity}, \S 3 \& \S 16, 3rd e    d., Butterworth-Heinemann, Oxford, UK (1986).

>>>>

@@

\bibitem{sulaiman2006}
N. Sulaiman, D. Marenduzzo, and J.M. Yeomans,
Lattice Boltzmann algorithm to simulate isotropic-nematic emulsions,
\textit{Phys. Rev. E} \textbf{74} 041708 (2006).

\bibitem{swift1996}
M.R. Swift, E. Orlandini, W.R. Osborn, and J.M. Yeomans,
Lattice Boltzmann simulation of liquid-gas and binary fluid systems,
\textit{Phys. Rev. E}, \textbf{54} 5041 (1996).

@@

\bibitem{Lyklema} J. Lyklema {\em Fundamentals of Interface and Colloid Science} Academic Press (1995).
\bibitem{Mafe} S. Maf\'e, J.A. Manzanares, J. Pellicer, {\textit J. Electroanal. Chem.} {\textbf 241}, 57-77 (1988).
\bibitem{Capuani} F. Capuani, I. Pagonabarraga, D. Frenkel, {\textit J. Chem. Phys.} {\textbf 121}, 973-986 (2004).
\bibitem{Rotenberg} B. Rotenberg, I. Pagonabarraga, D. Frenkel, {\textit Farad. Discuss.} {\textbf 144}, 223-243 (2010).
\bibitem{Landau-ED} L.D. Landau, E.M. Lifshitz, {\textit Electrodynamics of Continuous Media}, \S 15, 2nd ed., Pergamon Press, Oxford, UK (1984).
\bibitem{Melcher} J.R. Melcher, {\textit Continuum Electromechanics}, \S 3.10, MIT Press, Cambridge, MA, USA (1981).\\
downloadable from:\\
\url{http://ocw.mit.edu/ans7870/resources/melcher/resized/cem_811.pdf}
\bibitem{Landau-EL} L.D. Landau, E.M. Lifshitz, {\textit Theory of Elasticity}, \S 3 \& \S 16, 3rd ed., Butterworth-Heinemann, Oxford, UK (1986).
